Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-10-17 src/tex/ms.tex
Update ms.tex

% Section guide
In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
In Sec.~\ref{sec:Method}, we describe the modification we apply to the baseline \ac{GR} waveform, summarize our \ac{PE} configuration and outline our statistical methods for combining information from multiple events.
In Sec.~\ref{sec:Results}, we present our constraint on \ac{GW} amplitude birefringence, discussing individual events and the catalog collectively.
In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
\added{Additional details on the results for some notable events are presented in \ref{sec:Appendix}.}

@@

Equation \eqref{eq:waveform_modification} can be derived as the first-order effect in an expansion away from \ac{GR} under multiple frameworks.
In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudoscalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
Since it originates from a truncated series expansion,%
\footnote{Often the birefringent effect is written as an expansion in redshift $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance. Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.  Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the Universe and constrain this effect meaningfully.} %
Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, namely
\begin{equation}
    \left(\frac{\left|\kappa\right|}{\mathrm{Gpc}^{-1}}\right) \left(\frac{d_c}{\mathrm{Gpc}}\right) \left(\frac{f}{100\, \mathrm{Hz}}\right) < 1\,,
    \label{eq:small_exponent}
\end{equation}
recalling that $\kappa$ has dimensions of inverse length.
Otherwise, more frequency-dependent terms could enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
\begin{figure}
    \script{birefringence.py}
    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
    \caption{
        Illustration of amplitude birefringence. The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top). 
        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
        }
    \label{fig:birefringence}
\end{figure}

@@

\appendix

\section{EXTENDED RESULTS FOR NOTABLE EVENTS}
\label{sec:Appendix}

>>>>

% Section guide
In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
In Sec.~\ref{sec:Method}, we describe the modification we apply to the baseline \ac{GR} waveform, summarize our \ac{PE} configuration and outline our statistical methods for combining information from multiple events.
In Sec.~\ref{sec:Results}, we present our constraint on \ac{GW} amplitude birefringence, discussing individual events and the catalog collectively.
In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
\added{In the Appendix, we provide extended results and discussion for two notable events (GW170818 and GW190521).}

@@

Equation \eqref{eq:waveform_modification} can be derived as the first-order effect in an expansion away from \ac{GR} under multiple frameworks.
In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudoscalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
Since it originates from a truncated series expansion,%
\footnote{Often the birefringent effect is written as an expansion in redshift $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance. Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.  Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the Universe and constrain this effect meaningfully.} %
Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, namely
\begin{equation}
    \left(\frac{\left|\kappa\right|}{\mathrm{Gpc}^{-1}}\right) \left(\frac{d_c}{\mathrm{Gpc}}\right) \left(\frac{f}{100\, \mathrm{Hz}}\right) < 1\,,
    \label{eq:small_exponent}
\end{equation}
recalling that $\kappa$ has dimensions of inverse length.
Otherwise, more frequency-dependent terms could enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
\begin{figure}
    \script{birefringence.py}
    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
    \caption{
        Illustration of amplitude birefringence. The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
        However, if spacetime were birefringent \added{(BR)} following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top). 
        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
        }
    \label{fig:birefringence}
\end{figure}

@@

\appendix*

\section{Extended results for notable events}
\label{sec:Appendix}
